\section{Overview}
This section discusses the 5 stages of our system and their functionalities.

\subsection{Fetch Stage}
\begin{itemize}
    \item Responsible for fetching the next instruction.
    \item Can take two cycles in case of 32-bit instructions.
    \item Contains a branch prediction unit to determine the next address to be fetched in case of branching.
    \item Outputs the instruction bits into IF/ID Buffer.
    \item The current instruction is fetched at the first half of cycle, then the next PC value calculations are done in the second half.
\end{itemize}

\subsection{Decode Stage}
\begin{itemize}
    \item Responsible for decoding the instruction bits into control signals.
    \item Outputs the corresponding signals to ID/EX Buffer.
    \item Contains register file to output operand values and register-related operations.
    \item Determines the correct branch address in case of branching instructions by using Branch Address Unit.
    \item The control unit deduces the corresponding signals in the first half of cycle, then the register operations and branch address calculation are done in the second half of cycle.
\end{itemize}

\subsection{Execute Stage}
\begin{itemize}
    \item Responsible for ALU operations.
    \item Determines the correct ALU output and pass it with other signals to EX/M Buffer.
    \item The ALU operations and CCR update are done in the first half of cycle.
\end{itemize}

\subsection{Memory Stage}
\begin{itemize}
    \item Responsible for Data Memory IO.
    \item Memory read/write is done in the first half of cycle.
\end{itemize}

\subsection{Write-Back Stage}
\begin{itemize}
    \item Responsible for passing correct output values to the destination registers.
    \item Write back is done in the first half of cycle.
\end{itemize}

\section{IF/ID Buffer}

\subsection{Registers}
\begin{itemize}
    \item Instruction Register (32 bits)
    \item Next Address Register (32 bits)
    \item Incremented PC Register (32 bits)
    \item Hashed Address Register (4 bits)
    \item Interrupt Register (1 bit)
\end{itemize}

\subsection{Control Signals}
\begin{itemize}
    \item Flush: clear buffer (1 bit)
    \item Stall: freeze buffer (1 bit)
\end{itemize}

\section{ID/EX Buffer}

\subsection{Registers}
\begin{itemize}
    \item Operand Registers (2$X$32 bits)
    \item Destination Register (4 bits)
    \item OpCode Register (7 bits)
    \item R/W Register (2 bits)
\end{itemize}

\subsection{Control Signals}
\begin{itemize}
    \item Stall (IN): freeze buffer (1 bit)
    \item Destination Register (OUT) (4 bits)
    \item Output Values (OUT) (32 bits)
\end{itemize}

\section{EX/M Buffer}

\subsection{Registers}
\begin{itemize}
    \item ALUout Register (32 bits)
    \item MEM IN Register (32 bits)
    \item Opcode Register (7 bits)
    \item Destination Register (4 bits)
    \item R/W Register (2 bits)
\end{itemize}

\subsection{Control Signals}
\begin{itemize}
    \item Destination Register (OUT) (4 bits)
    \item Output Values (OUT) (32 bits)
\end{itemize}

\section{M/WB Buffer}

\subsection{Registers}
\begin{itemize}
    \item ALUout Register (32 bits)
    \item MEM OUT (32 bits)
    \item OpCode (7 bits)
    \item Destination Register (4 bits)
\end{itemize}
