\section{Completeness}
The processor is completely implemented except for \textbf{memory cache system}, which is already designed. However, we ran out of time trying to integrate it to our current implementation.

So, we reverted back to the two memories, \emph{Harvard} architecture. However, we included the cache system design in this report to share our thinking.

\section{Functionality}
The implemented processor, with \emph{Harvard} architecture, completely works with correct functionalities and hazard handling.

However, we still have the following issues:
\begin{itemize}
    \item Some instructions cause data hazards, that can't be completely solved by HDU. So, we had to use some software solutions to handle it, due to time constraints (Refer to \ref{sec:software}).
\end{itemize}

\section{General Notes}
\begin{itemize}
    \item In our implementation, the data memory is divided into \textbf{two parts}. In order to handle the 32-bit bus with the 16-bit width. \textbf{Even addresses} are written to one part and \textbf{odd addresses} are written to the other. The input address to the whole data memory is even, then the data memory receives this address and extract the corresponding odd address. Data memory is implemented this way, in order to be able to read or write an entire 32-bit entry in one cycle. 
    \item Fetch stage outputs a single \emph{NOP} instruction, in case of \emph{Interrupt}, \emph{CALL}, \emph{JZ} and \emph{JMP}, due to the delay of reading data (PC values or branch address) from register file.
    \item Flushing happens only to the fetch stage, in case of incorrect predicted branch address for \emph{JZ} instruction.
    \item The compiler should add an extra instruction \emph{END} to the end of the assembly program to mark the end of the code and terminate the simulator.
    \item We used \emph{ghdl} as a \emph{VHDL} compiler and \emph{GTKWave} as a simulator throughout the development process for fast development and debugging. We included our development setup, along with the main deliverables of the project. 
\end{itemize}